---
title: "Harvard Data Science Professional Capstone Project - MovieLens Quizzes"
author: "Uday Adusumilli"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# setup code
Create edx set, validation set (final hold-out test set)

step 1
Note: this process could take a couple of minutes
```{r}
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(data.table)) install.packages("data.table", repos = "http://cran.us.r-project.org")

library(tidyverse)
library(caret)
library(data.table)
```

The following libraries were needed, and they were not mentioned in the code provided to us

```{r}

library(dplyr)
library(tidyverse)
library(kableExtra)
library(tidyr)
library(stringr)
library(forcats)
library(ggplot2)
```


```{r}
# MovieLens 10M dataset:
# https://grouplens.org/datasets/movielens/10m/
# http://files.grouplens.org/datasets/movielens/ml-10m.zip

dl <- tempfile()
download.file("http://files.grouplens.org/datasets/movielens/ml-10m.zip", dl)

ratings <- fread(text = gsub("::", "\t", readLines(unzip(dl, "ml-10M100K/ratings.dat"))),
                 col.names = c("userId", "movieId", "rating", "timestamp"))

movies <- str_split_fixed(readLines(unzip(dl, "ml-10M100K/movies.dat")), "\\::", 3)
colnames(movies) <- c("movieId", "title", "genres")
```

the following block will depend on the r version
```{r}
# if using R 3.6 or earlier:
# movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(levels(movieId))[movieId],
#                                             title = as.character(title),
#                                             genres = as.character(genres))
# if using R 4.0 or later:
movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(movieId),
                                            title = as.character(title),
                                            genres = as.character(genres))

```

the following block is version independent
```{r}
movielens <- left_join(ratings, movies, by = "movieId")

# Validation set will be 10% of MovieLens data
set.seed(1, sample.kind="Rounding") # if using R 3.5 or earlier, use `set.seed(1)`
test_index <- createDataPartition(y = movielens$rating, times = 1, p = 0.1, list = FALSE)
edx <- movielens[-test_index,]
temp <- movielens[test_index,]

# Make sure userId and movieId in validation set are also in edx set
validation <- temp %>% 
      semi_join(edx, by = "movieId") %>%
      semi_join(edx, by = "userId")

# Add rows removed from validation set back into edx set
removed <- anti_join(temp, validation)
edx <- rbind(edx, removed)

rm(dl, ratings, movies, test_index, temp, movielens, removed)
```


# Actual Quizz Questions

### Q1

How many rows and columns are there in the edx dataset?


```{r}
dim(edx)
```


### Q2

How many zeros were given as ratings in the edx dataset?

```{r}
edx %>% filter(rating == 0) %>% tally()
```

How many threes were given as ratings in the edx dataset?

```{r}

edx %>% filter(rating == 3) %>% tally()
```

### Q3

How many different movies are in the edx dataset?

```{r}

n_distinct(edx$movieId)
```

### Q4

How many different users are in the edx dataset?

```{r}

n_distinct(edx$userId)
```

### Q5

How many movie ratings are in each of the following genres in the edx dataset?

```{r}

edx %>% separate_rows(genres, sep = "\\|") %>%
  group_by(genres) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```

### Q6

Which movie has the greatest number of ratings?

```{r}
edx %>% group_by(movieId, title) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```

### Q7

What are the five most given ratings in order from most to least?

```{r}

edx %>% group_by(rating) %>% summarize(count = n()) %>% top_n(5) %>%
  arrange(desc(count))  
```

### Q8

True or False: In general, half star ratings are less common than 
whole star ratings (e.g., there are fewer ratings of 3.5 than there 
are ratings of 3 or 4, etc.).

```{r}

edx %>%
  group_by(rating) %>%
  summarize(count = n()) %>%
  ggplot(aes(x = rating, y = count)) +
  geom_line()
```

# h1
## h2
### h3
#### h4
regular line